### 2016-05-19 transformation.tex
redraw figure 5
polish Section 4 and include 2 references

With the above discussions, we need to answer two questions before adopting the period enforcer algorithm. First, how can we derive the corresponding deferrable task set efficiently? Second, how should we perform the schedulability test on the corresponding deferrable task set correctly?

@@

The transformation from a segmented self-suspending task set to a corresponding deferrable task set was not discussed in \cite{Raj:suspension1991}. However, in our opinion, in general, such a transformation is not an easy problem. We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al. \cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity. The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.

@@

Based on these considerations, it appears that, at least for the simple example, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems. Deriving the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not an easy problem.  The method recently provided by Nelissen et al. \cite{ecrts15nelissen} can be used for this specific case to derive the exact $W_k^j$ if we assume that task $\tau_k$ has only the first $j$ computation segments and $j-1$ self-suspension intervals.  However, Nelissen et al. \cite{ecrts15nelissen} also showed that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem, in which calculating $W_k^j$ would need exponential time complexity if $j \geq 2$.
In particular, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.

In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98}, if over-approximations can be tolerated.

@@

Here, we explain the potential schedulability test that can be used for a corresponding deferrable task set. It may be seemingly correct at beginning that we do not have to consider the interference from the other deferrable tasks created from one segmented self-suspending task. For example, in the special task set considered in Section~\ref{sec:transformation-exponential}, one may argue that it is not necessary to consider the interference from the other computation segments of task $\tau_k$ while analyzing the feasibility of the last computation segment. However, this statement is incorrect.

Consider a task system consisting of $2$ tasks. Let $\tau_1$ denote a sporadic task without self-suspensions and parameters $C_1 = 2$ and $T_1=D_1=10$, and let $\tau_2$ denote a self-suspending task consisting of three segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$. That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}. One additional computation
segment is created to illustrate the potential concerns. This task set is schedulable by fixed-priority preemptive scheduling without any enforcement since the worst-case response time of the second computation segment is $10$ and the third computation segment will suffer from the interference by at most one job from task $\tau_1$. Moreover, by using the analysis in Section~\ref{sec:transformation-exponential}, we can conclude that
\begin{equation*}
  \rho_2^1 = 0, \rho_2^2 = 9, \rho_2^3 = 18.
\end{equation*}
If we do not consider the interference from the first two computation segments of task $\tau_2$, we can conclude that the last computation segment of task $\tau_2$ has worst-case deferrable time (release jitter) $18$ and the worst-case response time of a job of task $\tau_2$ is hence $18+C_1+C_2^3=21$.

However, the above statement is not correct. Let us consider this task set under control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
Figure \ref{fig:example-deferrable-incompatible} shows the resulting schedule for a periodic release pattern. This changes the schedule of the second job of task $\tau_2$. 
The second job of task $\tau_2$ (which is released at time $21$) is affected by period enforcement. The first segment of the second job arrives at time $a^1_{2,2} = 21$, is interfered for one time unit, and suspends at time $23$. The  second segment of the second job hence resumes only at time $a^2_{2,2} = 29$. Thus, we have
\begin{align*}
	ET_{2,2}^1 & = \max\left(0 + 21,\ \mathit{busy}(\tau_2, 21)\right) = 21  \text{ and }
\\
	ET_{2,2}^2 & = \max\left(9 + 21,\ \mathit{busy}(\tau_2, 29\right) ) = 30.
\end{align*}
According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $29$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 30$. However, at time $30$, the fourth job of $\tau_1$ is released. As a result, the second job of $\tau_2$ suffers from additional interference and misses its deadline at time $42$.

If we further allow task $\tau_1$ to be released sporadically, the fifth job of task $\tau_1$ can be released at time $41$ instead of time $40$ in Figure~\ref{fig:example-deferrable-incompatible}. For such a case, the response time of the second job of task $\tau_2$ is $23$. This means that the worst-case response time of the deferrable task created to represent the third computation segment of task $\tau_2$ must be at least $23-\rho_2^3=5$, which equals to $C_1 + C_2^1+C_2^2+C_2^3$.

Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments.

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.4]{../figures/deferrable-incompatible.pdf}
  \caption{}
  \label{fig:example-deferrable-incompatible}
\end{figure}

>>>>

With the above discussions, we need to answer two questions before adopting the period enforcer algorithm. First, how can we derive the corresponding deferrable task set efficiently? Second, how should we perform the schedulability test on the corresponding deferrable task set correctly? We will discuss the difficulty to derive the corresponding deferrable task set in Section~\ref{sec:transformation-exponential} and the potential pessimism for the schedulability test based on the deferrable task set in Section~\ref{sec:schedulability-test-deferrable}.

@@

The transformation from a segmented self-suspending task set to a corresponding deferrable task set was not discussed in \cite{Raj:suspension1991}. However, in our opinion, in general, such a transformation is not an easy problem. We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al. \cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity. The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.\footnote{Here, we refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}. The exact worst-case response time is to explore all the release patterns that satisfy the conditions in Lemma 2 in \cite{ecrts15nelissen} instead of the proposed MILP in \cite{ecrts15nelissen} (which only provides an upper bound). }

@@

Based on these considerations, it appears that, at least for the simple example, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems. Deriving the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not an easy problem.  The method recently provided by Nelissen et al. \cite{ecrts15nelissen} can be used for this specific case to derive the exact $W_k^j$ if we assume that task $\tau_k$ has only the first $j$ computation segments and $j-1$ self-suspension intervals.  However, Nelissen et al. \cite{ecrts15nelissen} also showed that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem, in which calculating $W_k^j$ would need exponential time complexity if $j \geq 2$.
In particular, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.\footnote{The schedulability test problem for such a case has been recently proved to be coNP-hard in the strong sense by Jian-Jia Chen in a report.}

In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, if over-approximations can be tolerated.

@@

Suppose that the corresponding deferrable task set can be derived precisely by ignoring the complexity issues raised in Section~\ref{sec:transformation-exponential}. The next question is whether there is any information loss, with respect to the schedulability test, by inspecting only the corresponding deferrable task set. We will demonstrate an example to explain the information loss.

Consider a task system consisting of $2$ tasks. Let $\tau_1$ denote a sporadic task without self-suspensions, in which $C_1 = 2$ and $T_1=D_1=10$. Let $\tau_2$ denote a self-suspending task consisting of three computation segments with parameters  $C_2^1 = 1$,  $S_2^1 = 6$, $C_2^2=1$, $S_2^2=8$, $C_2^3=1$, and $ T_2=D_2=21$. That is, this task set is pretty similar to the task set used in Section~\ref{sec:unschedulable}. One additional computation
segment is created to illustrate the potential concerns. This task set is schedulable by fixed-priority preemptive scheduling without any enforcement since the worst-case response time of the second computation segment is $10$ and the third computation segment will suffer from the interference by at most one job from task $\tau_1$. Moreover, by using the analysis in Section~\ref{sec:transformation-exponential}, we can conclude that
\begin{equation*}
  \rho_2^1 = 0, \rho_2^2 = 9, \rho_2^3 = 18.
\end{equation*}
Therefore, in the corresponding deferrable task set, there are $2$ sporadic tasks $\tau_1$ and $\tau_2^1$ (with execution time $1$ and period $21$), and two deferrable tasks $\tau_2^2$ (with deferrable length $9$, period $21$, and execution time $1$) and $\tau_2^3$ (with deferrable length $18$, period $21$, and execution time $1$).

Suppose that we are interested to analyze the worst-case response time of the deferrable task $\tau_2^3$ (in the corresponding deferrable task set) under fixed-priority scheduling. The question is \emph{whether we should ignore or should include the interference due to $\tau_2^1$ and $\tau_2^2$}.
Since the computation segments that are used to create $\tau_2^1, \tau_2^2$, and $\tau_2^3$ do not have any overlap in their available time to be executed, it may be seemingly correct at beginning that we do not have to consider the interference from $\tau_2^1$ and $\tau_2^2$ when analyzing the response time of the deferrable task $\tau_2^3$. However, this statement is in fact incorrect.

If we do not consider the interference from $\tau_2^1$ and $\tau_2^2$, we can conclude that the worst-case response time of $\tau_2^3$ is  $18+C_1+C_2^3=21$.
Let us consider this task set under the control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
Figure \ref{fig:example-deferrable-incompatible} shows the resulting schedule for a periodic release pattern. This changes the schedule of the second job of task $\tau_2$. 
The second job of task $\tau_2$ (which is released at time $21$) is affected by period enforcement. The first segment of the second job arrives at time $a^1_{2,2} = 21$, is interfered for one time unit, and suspends at time $23$. The  second segment of the second job hence resumes only at time $a^2_{2,2} = 29$. Thus, we have
\begin{align*}
	ET_{2,2}^1 & = \max\left(0 + 21,\ \mathit{busy}(\tau_2, 21)\right) = 21  \text{ and }
\\
	ET_{2,2}^2 & = \max\left(9 + 21,\ \mathit{busy}(\tau_2, 29\right) ) = 30.
\end{align*}
According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $29$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 30$. However, at time $30$, the fourth job of $\tau_1$ is released. As a result, the second job of $\tau_2$ suffers from additional interference and misses its deadline at time $42$.

If we further allow task $\tau_1$ to be released sporadically, the fifth job of task $\tau_1$ can be released at time $41$ instead of time $40$ in Figure~\ref{fig:example-deferrable-incompatible}. For such a case, the response time of the second job of task $\tau_2$ is $23$. This means that the worst-case response time of the deferrable task$\tau_2^3$ must be at least $23$, which equals to $18+C_1 + C_2^1+C_2^2+C_2^3$.

Therefore, for this specific example, we know that the schedulability test for the  corresponding deferrable task set has also to be pessimistic enough to consider the interferences from all the computation segments created from the same self-suspending sporadic task.

\begin{figure}[t]
  \centering  
  \includegraphics[scale=1]{../figures/deferrable-incompatible/deferrable-incompatible.pdf}
  \caption{An example for explaining why it is not safe to ignore the interference from the other computation segments in the same task.}
  \label{fig:example-deferrable-incompatible}
\end{figure}

### 2016-05-19 untitled.tex
take out unneeded macros (compilation trouble)
polish Section 4 and include 2 references

\section{Introduction} 
When real-time tasks suspend themselves (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time. A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95}. This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.

@@

\ifpaper
\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/not-ok-without-period-enforcer/not-ok.pdf}
  \caption{Example uniprocessor schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$. Tasks $\tau_1$ and $\tau_3$ consist of a single computation segment ($C_1^1 = C_3^1 = 3$); task $\tau_2$ consists of two computation and one suspension segment ($C_2^1 = 1$, $S_2^1 = 4$, $C_2^2 = 2$). Jobs of tasks $\tau_1$ and $\tau_3$ are released just as $\tau_2$ resumes from its self-suspension at time~$5$. Without period enforcement, task $\tau_3$ misses a deadline at time~$15$ because the second job of  task $\tau_2$ suspends only briefly (for one time unit rather than four).}
  \label{fig:not-ok-without-period-enforcement}
  \end{figure}
\fi

@@

\ifpaper
\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/ok-with-period-enforcer/ok.pdf}
  \caption{Example uniprocessor schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}. With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'').}
  \label{fig:ok-with-period-enforcement}
  \end{figure}
\fi

@@

\ifpaper
\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/example-original/example-original.pdf}
  \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals on a uniprocessor. Task $\tau_1$ has higher priority than task $\tau_2$.}
  \label{fig:example-original}
  \end{figure}
\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/example/example.pdf}
  \caption{An illustrative example demonstrating a deadline miss at time 22
    under the period enforcer algorithm. At time~19, $\tau_2$ resumes, but it remains ineligible to execute until time~20 when $\tau_1$ is released.}
\label{fig:example}  
\end{figure}
\fi

@@

>>>>

\section{Introduction} 
When real-time tasks suspend themselves (due to blocking I/O, lock contention, etc.), they defer a part of their execution to be processed at a later time. A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}. This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/not-ok-without-period-enforcer/not-ok.pdf}
  \caption{Example uniprocessor schedule (\emph{without} period enforcement) of three tasks $\tau_1 = $, $\tau_2$, and $\tau_3$ with periods $T_1 = T_2 = T_3 = 10$. Tasks $\tau_1$ and $\tau_3$ consist of a single computation segment ($C_1^1 = C_3^1 = 3$); task $\tau_2$ consists of two computation and one suspension segment ($C_2^1 = 1$, $S_2^1 = 4$, $C_2^2 = 2$). Jobs of tasks $\tau_1$ and $\tau_3$ are released just as $\tau_2$ resumes from its self-suspension at time~$5$. Without period enforcement, task $\tau_3$ misses a deadline at time~$15$ because the second job of  task $\tau_2$ suspends only briefly (for one time unit rather than four).}
  \label{fig:not-ok-without-period-enforcement}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/ok-with-period-enforcer/ok.pdf}
  \caption{Example uniprocessor schedule \emph{with} period enforcement assuming the same scenario as depicted in Figure~\ref{fig:not-ok-without-period-enforcement}. With period enforcement, task $\tau_3$ does not miss a deadline because task $\tau_2$'s second computation segment is delayed until it no longer imposes undue interference (i.e., it is prevented from resuming ``too soon'').}
  \label{fig:ok-with-period-enforcement}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/example-original/example-original.pdf}
  \caption{An illustrative example of the original self-suspending task set (without period enforcement) assuming periodic job arrivals on a uniprocessor. Task $\tau_1$ has higher priority than task $\tau_2$.}
  \label{fig:example-original}
  \end{figure}
\begin{figure}[t]
  \centering
  \includegraphics[scale=1]{../figures/example/example.pdf}
  \caption{An illustrative example demonstrating a deadline miss at time 22
    under the period enforcer algorithm. At time~19, $\tau_2$ resumes, but it remains ineligible to execute until time~20 when $\tau_1$ is released.}
\label{fig:example}  
\end{figure}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "LITES/LITES-Paper.tex"
%%% End:
